---
title: "Untitled"
author: "Patrick Schwarz"
date: "7 1 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold")
library(fastDummies)
library(AER)
library(stargazer)
library(lmtest)
library(equatiomatic)
library(ggplot2)
library(tidyverse)
library(knitr)
library(kableExtra)
library(hrbrthemes)

```

```{r}
vote <- read.csv("ESS8.csv", 
                 header = TRUE, 
                 sep    = ";", 
                 quote  = "",
                 dec    = ".", 
                 fill   = TRUE)
```

Table1: Variables in the data set

Variable                    Description          
---------                   ---------         
Party                       Name of the party the respondent voted last election
Trust in Politicians        Categorical response with 3 levels
Age                         Age of the respondent
Religiousness               Likert scale from 0-10 min to max
Gender                      Male or Female
Happiness                   Likert scale from 0-10 min to max
Health                      Health of the respondent from very good to very bad
Caring for Others           Ordinal in 6 steps from very much to not at all

```{r}
# Number of obs before removing rows
NumObs_before <- nrow(vote)

vote <- vote[ which(vote$trstplt   < 11),]

vote <- vote[ which(vote$happy   < 11),]
vote <- vote[ which(vote$prtvtbse  < 13),]
vote <- vote[ which(vote$health  < 6),]
vote <- vote[ which(vote$rlgdgr   < 11),]
vote <- vote[ which(vote$agea  < 91),]
vote <- vote[ which(vote$iphlppl   < 7),] #care
vote <- vote[ which(vote$imptrad   < 7),] #tradition


# Number of obs after removing rows
NumObs_after <- nrow(vote)

# Number removed 
NumObs_removed = NumObs_before-NumObs_after
NumObs_removed

```

```{r}

vote <-
  vote %>%
   rename(  Trustfulness = ppltrst,
            Helpfulness = pplhlp,
            Trust_politicians  = trstplt,
            Happiness = happy, 
            Health = health,
            Care_for_others = iphlppl ,
            Traditionalist = imptrad ,
            Hedonist = impfun,
            Age = agea,
            Party = prtvtbse)

```



```{r}
vote2 <-
  vote %>%
  mutate(
    Trust_politicians = recode(
      Trust_politicians,
      "0" = "less trusting",
      "1" = "less trusting",
      "2" = "less trusting",
      "3" = "less trusting",
      "4" = "less trusting",
      "5" = "neutral",
      "6" = "more trusting",
      "7" = "more trusting",
      "8" = "more trusting",
      "9" = "more trusting",
      "10" = "more trusting")) %>%
  mutate(
    Health = recode(
      Health,
      "1" = "Very good",
      "2" = "Good",
      "3" = "Fair",
      "4" = "Bad",
      "5" = "Very bad")) %>%
  mutate(
    Traditionalist = recode(
      Traditionalist,
      "1" = "Very much like me",
      "2" = "Like me",
      "3" = "Somewhat like me",
      "4" = "A little like me",
      "5" = "Not like me",
      "6" = "Not like me at all")) %>%
  mutate(
    Hedonist = recode(
      Hedonist,
      "1" = "Very much like me",
      "2" = "Like me",
      "3" = "Somewhat like me",
      "4" = "A little like me",
      "5" = "Not like me",
      "6" = "Not like me at all")) %>%
  mutate(
    Care_for_others = recode(
      Care_for_others,
      "1" = "Very much like me",
      "2" = "Like me",
      "3" = "Somewhat like me",
      "4" = "A little like me",
      "5" = "Not like me",
      "6" = "Not like me at all")) %>%
  mutate(
    Gender = recode(
      gndr,
      "1" = "Male",
      "2" = "Female",
      ))%>%
  mutate(
    Party = recode(
      Party,
      "1" = "Center",
      "2" = "Liberals",
      "3" = "Christian Democrats",
      "5" = "Conservatives",
      "4" = "Green Party",
      "6" = "Social Democrats",
      "7" = "Left Party",
      "8" = "Feminist Party",
      "9" = "Pirate Party",
      "10" = "Swedish Democrats",
      "11" = "Other"
      )) %>%
  mutate(Farbe = case_when(
           Party == "Center" ~ "#B3FFB3",
           Party == "Liberals" ~ "#99CCFF" ,
      Party == "Christian Democrats" ~ "#B0C4DE" ,
      Party ==  "Conservatives" ~ "#CCDDFF" ,
      Party == "Green Party" ~ "#CDE6AC" ,
      Party == "Social Democrats" ~ "#ed1b34" ,
      Party ==  "Left Party" ~ "#FF6E6E" ,
      Party ==  "Feminist Party" ~ "#FFBBBB" ,
      Party ==  "Pirate Party" ~ "#CC99CC" ,
      Party == "Swedish Democrats" ~ "#fedf09" ,
      Party == "Other" ~ "Grey" ))

# From the documentation of the variable prtvtbse
# Values and categories:
#  1	Centern (center party)
#  2	Folkpartiet liberalerna (liberals)
#  3	Kristdemokraterna (christian democrats)
#  4	Miljöpartiet de gröna (green party)
#  5	Moderata samlingspartiet (conservatives)
#  6	Socialdemokraterna (social democrates)
#  7	Vänsterpartiet (left party)
#  8	FI (Feministiskt initiativ) (feminist party)
#  9	Piratpartiet (pirate party)
# 10	Sverigedemokraterna (swedish democrats)
# 11	Annat parti (other party)
# 66	Not applicable
# 77	Refusal
# 88	Don't know
# 99	No answer
    
vote2$Trust_politicians <- factor(vote2$Trust_politicians, levels = c(
  "less trusting", "neutral","more trusting"), ordered=FALSE)

vote2$Health <- factor(vote2$Health, levels = c(
      "Very good",
      "Good",
      "Fair",
      "Bad",
      "Very bad"), ordered=FALSE)

vote2$Traditionalist <- factor(vote2$Traditionalist, levels = c(
 "Very much like me",
 "Like me",
 "Somewhat like me",
 "A little like me",
 "Not like me",
  "Not like me at all"), ordered=FALSE)

vote2$Hedonist <- factor(vote2$Hedonist, levels = c(
 "Very much like me",
 "Like me",
 "Somewhat like me",
 "A little like me",
 "Not like me",
  "Not like me at all"), ordered=FALSE)

vote2$Care_for_others <- factor(vote2$Care_for_others, levels = c(
 "Very much like me",
 "Like me",
 "Somewhat like me",
 "A little like me",
 "Not like me",
  "Not like me at all"), ordered=FALSE)

vote2$Gender <- factor(vote2$Gender)
vote2$Party <- factor(vote2$Party)

```

```{r}
fill_colors = as.character(vote2$Farbe)
names(fill_colors) = vote2$Party
```


```{r}
task1 <- read.csv('votesT.csv',header = TRUE, 
                 sep    = ";", 
                 quote  = "",
                 dec    = ",", 
                 fill   = TRUE)

colnames(task1) <- c('party', 'ESS Sample', '2014 Election', 'delta')

task1 %>%
  pivot_longer(c('ESS Sample', '2014 Election'), names_to = "year", values_to = "Votes") %>%
  select(-starts_with("delta")) %>%
  ggplot()+
  geom_bar(aes(x = fct_infreq(party), y=Votes, fill = party), stat = 'identity')+
  facet_grid(vars(year)) +
  theme(axis.text.x = element_text(angle = 35, vjust = 1, hjust=1, size = 8)) +
  geom_text(aes( x = party, y=Votes, label =Votes), size = 3) +
  guides(fill=FALSE) +
  xlab('Parties') + ylab ("Vote Percentage") +  
  scale_fill_manual(values = fill_colors)
```



```{r}

  
p <- vote2 %>%
  ggplot( aes(x=Party, y=Age, fill=Party)) + 
    geom_boxplot() +
    xlab("class") +
    theme(legend.position="none") +
    xlab("") +  scale_fill_manual(values = fill_colors) + coord_flip() + xlab("Party")
  
p
```




Force discrete palette
```{r}
ggplot(data=vote2, aes(x=rlgdgr, group=Party, fill=Party)) +
    geom_density() +
    theme_ipsum() +
    facet_wrap(~Party) +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      axis.ticks.x=element_blank()) +  
  scale_fill_manual(values = fill_colors) +
  
  theme(axis.text.x = element_text(size = 6)) +
    theme(axis.text.y = element_text(size = 8)) +
  xlab('Religiousness from not at all (0) to very much (10)')+ylab("Distribution of Answers")
 
```

```{r}
library(productplots)

prodplot(vote2, ~Traditionalist + Gender + stacked()) + aes(fill = Gender) +
    theme(legend.position="none")
sda <- prodplot(vote2, ~Trust_politicians + Party + stacked()) + labs(y="Trust in Politicians")+ aes(fill = Party) + coord_flip() +   
  scale_fill_manual(values = fill_colors)+
    theme(legend.position="none") 
set_y_axis_label <- function(pl, new_label) {
  pl$scales$scales[[1]]$name <- new_label
  pl
}
set_y_axis_label(sda, "Trust in Politicians")
sda
```
```{r}
vote3 <-
  vote2 %>%
    filter( Party != "Other") %>%
    filter( Party != "Pirate Party") %>%
    filter( Party != "Feminist Party") %>%
    mutate(Farbe = case_when(
           Party == "Center" ~ "#B3FFB3",
           Party == "Liberals" ~ "#99CCFF" ,
      Party == "Christian Democrats" ~ "#B0C4DE" ,
      Party ==  "Conservatives" ~ "#CCDDFF" ,
      Party == "Green Party" ~ "#CDE6AC" ,
      Party == "Social Democrats" ~ "#ed1b34" ,
      Party ==  "Left Party" ~ "#FF6E6E" ,
      Party == "Swedish Democrats" ~ "#fedf09"  ))

fill_colors2 = as.character(vote3$Farbe) 
names(fill_colors2) = vote3$Party 
tibble( fill_colors2)

vote3 %>%
    ggplot(aes(y = Happiness, x = Health, group = Party,  colour = fill_colors2)) +
        geom_point(position=position_jitter(h=1, w=1),
                 shape = 21, alpha = 0.5, size = 2) +
        facet_grid(vars(Gender), vars(Party))   +
theme(strip.text = element_text(angle =33,   size=6)) +
      
      theme(axis.text.x = element_text(angle = 90 ,size = 8)) +
        theme(axis.text.y = element_text(size = 8)) +
        theme(legend.position="none")
 
```




```{r}
vote2 %>%
  ggplot(aes(Party, fill = Care_for_others)) +
    geom_bar(position="fill") +
  ylab("Proportion") +
      
      theme(axis.text.x = element_text(size = 8)) +
  coord_flip() +
  scale_fill_ordinal(name = "Caring for Others")
```


```{r}
    require(extrafont)
    # need only do this once!
    font_import(pattern="[A/a]rial", prompt=FALSE)
    require(ggplot2)
    # extending the help file example
    df <- data.frame(gp = factor(rep(letters[1:3], each = 10)), y = rnorm(30))
    ds <- plyr::ddply(df, "gp", plyr::summarise, mean = mean(y), sd = sd(y))
    plotobj <- ggplot(df, aes(gp, y)) +
      geom_point() +
      geom_point(data = ds, aes(y = mean), colour = 'red', size = 3) + 
      theme(text=element_text(size=16))
    print(plotobj)
```

